R Notebook for conducting Minimum Detectable Change (MDC) Analysis to design a monitoring strategy for the Upper Yaquina River TMDL Implementation Plan

### 2024-02-15 msc.Rmd
no major changes made. may be update to project files after closing R

>>>>

## MDC function for Varying N

I used ChatGPT 4.0 to get a starting point for the function

```{r}
calculateMDCWithAdditionalSamples <- function(model, additional_samples, alpha, power) {
  # Extract the standard error of the slope (SE_b) directly from the model
  SE_b <- summary(model)$coefficients[2, "Std. Error"]

  # Calculate the total number of samples (existing plus additional)
  initial_n <- length(model$model[[1]]) # Assuming the first column in model data is the response variable
  total_n <- initial_n + additional_samples

  # Calculate degrees of freedom for the total sample size
  df <- total_n - 2

  # Calculate the critical t-values for the given alpha and power
  t_alpha <- qt(1 - alpha/2, df)
  t_beta <- qt(power, df)

  # Calculate the MDC using the SE_b from the model
  # Note: This calculation assumes SE_b remains constant with additional samples, which might not be accurate
  MDC <- (t_alpha + t_beta) * SE_b

  return(MDC)
}
```

No I tried the function for station 12301, which had the largest MDC

df_cur <- df_log10 %>% filter(STATION == "12301") 
str(df_cur)

model_cur <- lm(log10_EC_val ~ date, df_cur)

initial_n <- length(model_cur$model[[1]])
additional_samples <- 36

MDC <- calculateMDCWithAdditionalSamples(model_cur,additional_samples, 
                                         alpha = 0.05, power = 0.9)

print(paste("MDC for", initial_n + additional_samples, "total samples:", MDC))
